---
title: "Different Models"
output: html_notebook
---

This document demonstrates the different types of model which exist and their varying effectiveness.

```{r message=FALSE}
library(tidyverse)
```


Creating a Dataset
==================

We're going to make a dataset based off the IMDB film database.  We're using the database downloaded on 27th March 2023.  It comes in several parts which we're going to import and merge.

First we'll get the basic film information.

```{r}
read_delim("Datasets/IMDB/title.basics.tsv") -> films
```

We'll just use the films, not TV series etc.

```{r}
films %>%
    filter(titleType == "movie") -> films
```


Clean up some columns we're not going to use.

```{r}
films %>%
  select (-primaryTitle, -endYear) %>%
  mutate(runtimeMinutes = as.integer(runtimeMinutes)) %>%
  filter(!is.na(runtimeMinutes)) %>%
  filter(!endsWith(genres,"N")) -> films
```

We need to split the genres into separate columns.  First we'll extract the full list of genres

```{r}
films %>%
  distinct(genres) %>%
  arrange(desc(str_count(genres,","))) %>%
  separate_wider_delim(genres,delim = ",",names_sep = ":", too_few = "align_start") %>%
  pivot_longer(
    cols=everything(),
    names_to="temp",
    values_to="genre"
  ) %>%
  distinct(genre) %>%
  filter(!is.na(genre)) %>%
  pull(genre) -> genres
```

Now we'll make a column for each of them.

```{r}
sapply(genres,function(x) films[[paste0("genre_",x)]] <<- as.numeric(str_detect(films$genres,x)))
```

Now we can lose the genres column

```{r}
films %>%
  select(-genres) -> films
```

Next we want to add information about the director.  We need the director table and the names.

```{r}
read_delim("Datasets/IMDB/title.crew.tsv") -> director
read_delim("Datasets/IMDB/name.basics.tsv") -> names

director %>%
  select(tconst, directors) %>%
  mutate(nconst=str_replace(directors,",.*","")) %>%
  select(-directors) %>%
  left_join(
    names %>%select(nconst, primaryName, birthYear)
  ) %>%
  right_join(films) %>%
  select(-nconst) %>%
  rename(director = primaryName, director_birthYear = birthYear) -> films

rm(director)
```


We want to know how many films each director has directed at the point of each films release.

```{r}
films %>%
  mutate(startYear = as.integer(startYear)) %>%
  filter(!is.na(startYear)) %>%
  mutate(director_birthYear = as.integer(director_birthYear)) %>%
  filter(!is.na(director_birthYear)) %>%
  rename(year_of_release = startYear) %>%
  group_by(director) %>%
  arrange(year_of_release) %>%
  mutate(directors_film = 1:n()) %>%
  ungroup() %>%
  mutate(directors_age = year_of_release - director_birthYear) -> films

```


