---
title: "Development Gene Expression"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

This shows the use of various model types to try to predict whether genes are development genes based on their gene expression.

The data comes from GSE1133 and the developmental genes come from GO:003250

```{r}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
```

Load the Data
=============

```{r}
read_delim("development_gene_expression.txt") -> data

# The predicted variable needs to be factor
data %>%
  mutate(Development=factor(Development)) %>%
  select(Development,everything()) -> data

# We want to randomly shuffle the rows so there is no structure
set.seed(123)
data %>%
  sample_frac() -> data

head(data)

```

We remove the gene names since we're not using those and then split the rest into test/training

```{r}
data %>%
  select(-gene) %>%
  initial_split(prop=0.8) -> split_data

split_data
```


Create a model
==============

Decision Tree
-------------

We can now build a model - we'll start with a simple decision tree

```{r}
decision_tree(tree_depth = 10) %>%
  set_mode("classification") %>%
  set_engine("rpart") -> model

model %>% translate()
```

```{r}
model %>%
  fit(Development ~ ., data=training(split_data)) -> model_fit

model_fit
```

### Train the model

```{r}
model_fit %>%
  predict(new_data=training(split_data)) %>%
  bind_cols(training(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count()
```


### Test the model

#### Original Data

```{r}
model_fit %>%
  predict(new_data=training(split_data)) %>%
  bind_cols(training(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() 
```

```{r}
model_fit %>%
  predict(new_data=training(split_data)) %>%
  bind_cols(training(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() %>%
  mutate(
    correct = .pred_class==Development
  ) %>%
  group_by(correct) %>%
  summarise(
    n=sum(n)
  )
```

#### New Data

```{r}
model_fit %>%
  predict(new_data=testing(split_data)) %>%
  bind_cols(testing(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count()
```

```{r}
model_fit %>%
  predict(new_data=testing(split_data)) %>%
  bind_cols(testing(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() %>%
  mutate(
    correct = .pred_class==Development
  ) %>%
  group_by(correct) %>%
  summarise(
    n=sum(n)
  )
```


Random Forest
-------------

```{r}
rand_forest(trees=100, min_n=5) %>%
  set_mode("classification") %>%
  set_engine("ranger") -> model

model %>% translate()
```

### Train the model

```{r}
model %>%
  fit(Development ~ ., data=training(split_data)) -> model_fit

model_fit
```

### Test the model

#### Original Data

```{r}
model_fit %>%
  predict(new_data=training(split_data)) %>%
  bind_cols(training(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() 
```

```{r}
model_fit %>%
  predict(new_data=training(split_data)) %>%
  bind_cols(training(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() %>%
  mutate(
    correct = .pred_class==Development
  ) %>%
  group_by(correct) %>%
  summarise(
    n=sum(n)
  )
```

#### New Data

```{r}
model_fit %>%
  predict(new_data=testing(split_data)) %>%
  bind_cols(testing(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count()
```

```{r}
model_fit %>%
  predict(new_data=testing(split_data)) %>%
  bind_cols(testing(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() %>%
  mutate(
    correct = .pred_class==Development
  ) %>%
  group_by(correct) %>%
  summarise(
    n=sum(n)
  )
```


Nearest Neighbour
-----------------

KNN model

```{r}
nearest_neighbor(neighbors = 20, weight_func = "triangular") %>%
  set_mode("classification") %>%
  set_engine("kknn") -> model

model %>% translate()
```

### Train the model

```{r}
model %>%
  fit(Development ~ ., data=training(split_data)) -> model_fit

model_fit
```

### Test the model

#### Original Data

```{r}
model_fit %>%
  predict(new_data=training(split_data)) %>%
  bind_cols(training(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() 
```

```{r}
model_fit %>%
  predict(new_data=training(split_data)) %>%
  bind_cols(training(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() %>%
  mutate(
    correct = .pred_class==Development
  ) %>%
  group_by(correct) %>%
  summarise(
    n=sum(n)
  )
```

#### New Data

```{r}
model_fit %>%
  predict(new_data=testing(split_data)) %>%
  bind_cols(testing(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count()
```

```{r}
model_fit %>%
  predict(new_data=testing(split_data)) %>%
  bind_cols(testing(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() %>%
  mutate(
    correct = .pred_class==Development
  ) %>%
  group_by(correct) %>%
  summarise(
    n=sum(n)
  )
```

Neural Net
----------


```{r}
mlp(epochs = 1000, hidden_units = 200, penalty = 0.01, learn_rate = 0.01) %>% 
  set_engine("brulee", validation = 0) %>% 
  set_mode("classification") -> model

model %>% translate()
```

### Train the model

```{r}
model %>%
  fit(Development ~ ., data=training(split_data)) -> model_fit

model_fit
```

### Test the model

#### Original Data

```{r}
model_fit %>%
  predict(new_data=training(split_data)) %>%
  bind_cols(training(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() 
```

```{r}
model_fit %>%
  predict(new_data=training(split_data)) %>%
  bind_cols(training(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() %>%
  mutate(
    correct = .pred_class==Development
  ) %>%
  group_by(correct) %>%
  summarise(
    n=sum(n)
  )
```

#### New Data

```{r}
model_fit %>%
  predict(new_data=testing(split_data)) %>%
  bind_cols(testing(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count()
```

```{r}
model_fit %>%
  predict(new_data=testing(split_data)) %>%
  bind_cols(testing(split_data)) %>%
  group_by(.pred_class, Development) %>%
  count() %>%
  mutate(
    correct = .pred_class==Development
  ) %>%
  group_by(correct) %>%
  summarise(
    n=sum(n)
  )
```


